Human Judgement Forecasting of COVID-19 in the UK

### 2023-03-28 manuscript/manuscript.tex

Mean scores for the four forecasting approaches were close to each other, with a mean WIS for two-week-ahead-forecasts of incident cases on the natural scale (log scale) of 38k (0.25) for the Hub-ensemble, 39k (0.24) for the direct crowd forecasts, 40k (0.22) for the combined crowd forecast and 46k (0.23) for the $R_t$ forecasts (see Table \ref{tab:scores} and Figure \ref{fig:performance}). Differences in mean scores between the four were small compared to the standard deviations of the scores within each forecasting approach on the natural (56k, 67k, 69k, 75k) and on the log scale (0.22, 0.27, 0.25, 0.24). 
% Could add a sentence that scores were not significant using a paired wilcoxon mann whitney test. 
On the natural scale, the Forecast Hub ensemble was ahead of the crowd forecasts (WIS two weeks ahead: 38k for the Hub ensemble and 40k for the combined human ensemble), with the difference gradually increasing with increasing forecast horizon (see Figure \ref{fig:performance}). On the log scale, the Forecast Hub ensemble was slightly worse than all crowd forecasting approaches at a two week forecast horizon, but again slightly ahead for forecasts four weeks into the future.

On the natural scale, the WIS as a measure of the absolute distance between forecast and observation increases or decreases with the magnitude of the forecast target \cite{bosseTransformationForecastsEvaluating2023, bracherEvaluatingEpidemicForecasts2021}. Average scores were therefore dominated by performance around the peak when cases were highest, in particular by forecasts made on the 19th of July for the 31st of July (see Figure \ref{fig:forecasts-scores}). $R_t$ forecasts tended to be higher than both the direct forecasts and the Forecast Hub ensemble, especially around the peak, and therefore showed the poorest performance in terms of WIS on the natural scale. When looking at WIS values on the log scale,
%(that is, when log-transforming forecasts and observations prior to applying the WIS), 
scores were more equally distributed across the study period and more weight is given to forecasts in June and July which underpredicted the extent to which case number would rise. Similarly to what was observed previously for case forecasts \citep{bosseComparingHumanModelbased2022, sherrattPredictivePerformanceMultimodel2022a}, all forecasting approaches exhibited underdispersion, meaning that forecasts on average were too narrow and not uncertain enough, with empirical coverage below nominal coverage. Underdispersion was most severe for the $R_t$ forecasts.

@@

Overall, performance for two-week-ahead-forecasts of incident deaths was comparable for the Hub-ensemble, the direct crowd forecasts and the combined forecasts, while performance for the $R_t$ forecast was significantly worse. On the natural scale (log scale), mean WIS values were 38 (0.22) for the Hub ensemble, 34 (0.27) for the direct forecasts, 40 (0.22) for the combined crowd forecast, and 80 (0.25) for the $R_t$ forecasts (Table \ref{tab:scores}). 
With the exception of the $R_t$ crowd forecasts, differences between forecasting approaches were again small compared to 
standard deviations of
the scores within each forecasting approach on the natural scale (27, 31, 42, 73) and on the log scale (0.036, 0.081, 0.073, 0.13). 
%On the log scale, direct human forecasts, the combined crowd ensemble and the Forecast Hub ensemble showed comparable performance, while $R_t$ forecasts performed again significantly worse.

On the natural scale, direct forecasts had a better WIS than the Hub ensemble and the other two human forecast ensembles for a two-week-ahead horizon (direct: 34, Hub ensemble: 38, combined crowd: 40, $R_t$: 80), while the Hub ensemble was again slightly ahead for a four week-ahead horizon (direct: 88, Hub ensemble: 85, combined crowd: 98, $R_t$: 150, see Figure \ref{fig:performance} and Table \ref{tab:scores-4}). The combined crowd forecast (containing both the direct and the $R_t$ forecasts) overall performed worse than the direct crowd forecast on the natural scale, but much better than the $R_t$ forecasts. On the log scale, the combined crowd forecasts even performed best among the four approaches.

From June to the end of July, $R_t$ forecasts overpredicted deaths and were noticeable higher than other forecasts, whereas in August, $R_t$ forecasts underpredicted deaths and were substantially lower than other forecasts (see Figure \ref{fig:forecasts-scores}). On the natural scale, the highest scores were concentrated towards the end of the study period, where death incidences were highest, while on the log scale, scores were more evenly distributed across the study period. $R_t$ forecasts show strong underdispersion with empirical coverage of the 50\% prediction intervals being close to zero and smaller than 50\% for the 90\% prediction intervals (see Figure \ref{fig:performance} and Table \ref{tab:scores}). The Forecast Hub ensemble was overdispersed with empirical coverage exceeding nominal coverage, and for the direct human forecasts, empirical coverage was close to nominal coverage.

@@

For cases, where participants could observe the case forecast implied by their $R_t$ forecast, predictive performance was similar between corresponding direct and $R_t$ forecasts for most forecasters who had submitted both (see Figure \ref{fig:comparison-direct-rt-individual}). For deaths, where forecasters could not see the incidence forecast implied by their $R_t$ forecast or manually adjust the case fatality rate, performance of the $R_t$ forecasts was significantly worse than the corresponding direct forecasts for most forecasters. This suggests that the convolution from latent infections implied by the $R_t$ forecasts to deaths did not capture the relationship between cases and deaths accurately in this instance despite performing well in other settings \citep{meakinComparativeAssessmentMethods2022}. This may be related to changes in the age decomposition of cases and the change in the case fatality ratio (CFR) over the course of the study period due to the rise of the Delta variant starting in May 2021 in the UK. This change was not accounted for in the simple model used and the CFR could not be adjusted by forecasters.

@@

>>>>

All forecasting approaches exhibited underdispersion when predicting cases, meaning that forecasts on average were too narrow and not uncertain enough. Empirical coverage was below nominal coverage for all forecasts and forecast horizons. 
For 50\% prediction intervals, empirical coverage was worst for the direct crowd forecasts (0.31), best for the $R_t$ forecasts (0.46) and in between for the Hub ensemble and the crowd ensemble (both 0.38, see Table \ref{tab:scores}). For 90\% prediction intervals, coverage was worst for the $R_t$ forecasts (0.62) and slightly better for the other approaches (all 0.69). Coverage for all forecasts deteriorated further with increasing forecast horizon (see Figure \ref{fig:performance}). 
%This underdisperion for case forecasts is in line with what was observed previously for case forecasts \citep{bosseComparingHumanModelbased2022, sherrattPredictivePerformanceMultimodel2022a}, all forecasting approaches exhibited underdispersion,

In terms of WIS on the log scale, all human forecasting approaches outperformed the Forecast Hub ensemble for two week ahead forecasts of cases (see Figure \ref{fig:performance}). WIS values relative to the Hub ensemble (=1) were 0.91 for the combined crowd ensemble, 0.96 for the direct crowd forecasts and 0.93 for the $R_t$ forecasts (see Table \ref{tab:scores}). 
In terms of WIS on the natural scale, however, the Hub ensemble was ahead of all human forecasting approaches. Relative WIS values on the natural scale for two week ahead forecasts were 1.05 for the combined crowd ensemble, 1.03 for the direct crowd forecasts and 1.2 for the $R_t$ forecasts.

Performance of the Hub ensemble relative to the human forecasting approaches improved with increasing forecast horizon (see Figure \ref{fig:performance}). For a four-week-ahead forecast horizon, the Hub ensemble outperformed all other approaches both on the log scale (rel. WIS values the human forecasts of 1.02, 1.05, 1.06) and on the natural scale (rel. WIS values of 1.21, 1.25, 1.3) (see Table \ref{tab:scores-4}).

In terms of relative model ranks for two week ahead forecasts, the Hub ensemble and the $R_t$ show a higher variance than the combined crowd ensemble and the direct forecasts (See Figure \ref{fig:performance-ranks}). Both the Hub ensemble and the $R_t$ forecast are most often in the first place, but also most often in the last place. The direct forecasts places relatively equally in places 1-4. The crowd ensemble never places fourth, but also has the lowest number of first places. Model ranks only change marginally when switching between the log and the natural scale.

When comparing WIS values on the log scale with those on the natural scale, scores were more equally distributed across the study period on the log scale and more weight was given to forecasts in June and July which underpredicted the extent to which case number would rise (see Figure \ref{fig:forecasts-scores}).
On the natural scale, the WIS as a measure of the absolute distance between forecast and observation increased or decreased with the magnitude of the forecast target \cite{bosseTransformationForecastsEvaluating2023, bracherEvaluatingEpidemicForecasts2021}. Average scores were therefore dominated by performance around the peak when cases were highest, in particular by forecasts made on the 19th of July for the 31st of July (see Figure \ref{fig:forecasts-scores}). Switching between scores on the log and on the natural scale had the strongest effect on the $R_t$ forecasts, which had a relative WIS value of 0.96 on the log scale and 1.2 on the natural scale. The $R_t$ forecasts tended to be higher than both the direct forecasts and the Forecast Hub ensemble, especially around the peak and were most severely penalised for overprediction (see Figure \ref{fig:performance}A). On the log scale, underprediction played a larger role, favouring the $R_t$ forecasts (see Figure \ref{fig:performance}C).

\begin{figure*}
\centering
\includegraphics[width=0.99\textwidth]{../output/figures/performance-ranks.png}
\caption{\bf{Performance ranks}....}
\label{fig:performance-ranks}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.99\textwidth]{../output/figures/performance-ranks-4.png}
\caption{\bf{Performance ranks}....}
\label{fig:performance-ranks-4}
\end{figure*}

@@

All forecasting approaches except the $R_t$ forecasts showed higher empirical coverage for deaths than for cases (see Figure \ref{fig:performance}). For 50\% prediction intervals, the Hub ensemble exceeded the nominal coverage noticeably (0.77) (see Table \ref{tab:scores}). $R_t$ forecasts failed to get close to nominal coverage (0.15), while the combined crowd ensemble and the direct forecasts had empirical coverage close to nominal coverage (both 0.54). For 90\% prediction intervals, the Hub ensemble again exceeded nominal coverage and covered all observations (1) while the $R_t$ forecasts again failed to get close to nominal coverage (0.46). The crowd ensemble exhibited a some underdispersion (0.77) while the direct forcecasts almost reached nominal coverage for two week ahead forecasts of deaths (0.85).

In terms of WIS on the log scale for two week ahead predictions of deaths, the combined crowd ensemble (0.07) and the direct crowd forecasts (0.99) were slightly ahead of the Hub ensemble, while the $R_t$ forecasts performed noticeably worse (1.98) (see Figure \ref{fig:performance} and Table \ref{tab:scores}). Interestingly, combining the $R_t$ forecasts and the direct forecasts led to an ensemble that performed better than either of them. 
In terms of WIS on the natural scale, only the direct forecasts (0.89) performed better for two week ahead death predictions than the Hub ensemble, while the combined crowd ensemble performed a bit worse (1.06) and the $R_t$ forecasts again noticeably worse (2.1).

% Maybe say something about increasing forecast horizons. But the pattern is unclear.

In terms of relative model ranks for two week ahead death forecasts, the $R_t$ forecasts took the fourth place most often, while the direct forecasts placed first most often (see Figure \ref{fig:performance-ranks)}. Again, the crowd ensemble never placed fourth.

When comparing scores on the log and on the natural scale, scores on the log scale are again more evenly distributed across the study period. On the natural scale, high scores were concentrated around the end of the study period, were death incidences were highest (see Figure \ref{fig:forecasts-scores}).

@@

For cases, where participants could observe the case forecast implied by their $R_t$ forecast, predictive performance was similar between corresponding direct and $R_t$ forecasts for most forecasters who had submitted both (see Figure \ref{fig:comparison-direct-rt-individual}). For deaths, where forecasters could not see the incidence forecast implied by their $R_t$ forecast or manually adjust the case fatality rate, performance of the $R_t$ forecasts was significantly worse. From June to the end of July, $R_t$ forecasts overpredicted deaths and were noticeable higher than other forecasts, whereas in August, $R_t$ forecasts underpredicted deaths and were substantially lower than other forecasts (see Figure \ref{fig:forecasts-scores}). In particular, $R_t$ forecasts for deaths were worse than the corresponding direct death forecasts for most forecasters (see Figure \ref{fig:comparison-direct-rt-individual}). Figure \ref{fig:comparison-direct-rt-individual} suggests that changing from the direct forecasting method to $R_t$ forecasting for cases may have improved scores for better forecasters and decreased scores for worse forecasters, although sample sizes are too small to draw definitive conclusions.

%This suggests that the convolution from latent infections implied by the $R_t$ forecasts to deaths did not capture the relationship between cases and deaths accurately in this instance despite performing well in other settings \citep{meakinComparativeAssessmentMethods2022}. This may be related to changes in the age decomposition of cases and the change in the case fatality ratio (CFR) over the course of the study period due to the rise of the Delta variant starting in May 2021 in the UK. This change was not accounted for in the simple model used and the CFR could not be adjusted by forecasters.

@@

%This underdisperion for case forecasts is in line with what was observed previously for case forecasts \citep{bosseComparingHumanModelbased2022, sherrattPredictivePerformanceMultimodel2022a}, all forecasting approaches exhibited underdispersion,
